---
title: "project2_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{project2_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(project2)
```

```{r}
library(here)
library(ggplot2)
library(tidytuesdayR)
library(dplyr)
library(lubridate)
library(forcats)
```

## Demonstrate fn_cos()  


```{r}
x_lab=seq(0,10, by=0.1)
y_lab=cos(x_lab)
k_1=fn_cos(x_lab, 1)
k_3=fn_cos(x_lab, 3)
k_5=fn_cos(x_lab, 5)
k_7=fn_cos(x_lab, 7)
k_9=fn_cos(x_lab, 9)
cos_data=data.frame(x_lab, y_lab, k_1, k_3, k_5, k_7, k_9)
ggplot(cos_data)+
  geom_point(aes(x=x_lab, y=y_lab))+
  geom_line(aes(x=x_lab, y=k_1, color="k_1"))+
  geom_line(aes(x=x_lab, y=k_3, color="k_3"))+
  geom_line(aes(x=x_lab, y=k_5, color="k_5"))+
  geom_line(aes(x=x_lab, y=k_7, color="k_7"))+
  geom_line(aes(x=x_lab, y=k_9, color="k_9"))+
  scale_color_manual(values = c("k_1" = "red", "k_3" = "orange", "k_5" = "yellow", "k_7" = "green", "k_9" = "blue"))+
  coord_cartesian(ylim = c(1, -10),)

```

When k increase, the estimation is more accurate. For small k, it become more inaccurate when x
become larger.

## Demonstrate fun_sin()  

```{r}
sin_x_lab=seq(0,10, by=0.1)
sin_y_lab=sin(sin_x_lab)
sin_k_1=fun_sin(sin_x_lab, 1)
sin_k_3=fun_sin(sin_x_lab, 3)
sin_k_5=fun_sin(sin_x_lab, 5)
sin_k_7=fun_sin(sin_x_lab, 7)
sin_k_9=fun_sin(sin_x_lab, 9)
sin_data=data.frame(sin_x_lab, sin_y_lab, sin_k_1, sin_k_3, sin_k_5, sin_k_7, sin_k_9)
ggplot(cos_data)+
  geom_point(aes(x=sin_x_lab, y=sin_y_lab))+
  geom_line(aes(x=sin_x_lab, y=sin_k_1, color="k_1"))+
  geom_line(aes(x=sin_x_lab, y=sin_k_3, color="k_3"))+
  geom_line(aes(x=sin_x_lab, y=sin_k_5, color="k_5"))+
  geom_line(aes(x=sin_x_lab, y=sin_k_7, color="k_7"))+
  geom_line(aes(x=sin_x_lab, y=sin_k_9, color="k_9"))+
  scale_color_manual(values = c("k_1" = "red", "k_3" = "orange", "k_5" = "yellow", "k_7" = "green", "k_9" = "blue"))+
  coord_cartesian(ylim = c(1, -10),)

```

## Demonstrate calculate_CI()  

The dataset is from tidytuesday, 

The data dictionary is https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-02-07/readme.md#data-dictionary:~:text=07/big_tech_companies.csv%27)-,Data,-Dictionary.

In this part, I would like to explore the open, close price and rate of return of 14 tech companies from 2010-2023.  Especially, take closer look for the IBM, which seems an obvious decrease, and I also did the confidence interval to see whether the rate of return is significant different than 0. The goal of this part is to see the practical significant vs the statistical significant. 

```{r}

if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

files = c("tuesdata_big_tech_stock_prices.RDS","tuesdata_big_tech_companies.RDS")
if (any(!file.exists(here("data", files)))) {
    tuesdata = tidytuesdayR::tt_load('2023-02-07')
    big_tech_stock_prices = tuesdata$big_tech_stock_prices
    big_tech_companies = tuesdata$big_tech_companies

    # save the files to RDS objects
    saveRDS(tuesdata$big_tech_stock_prices, file = here("data", "tuesdata_big_tech_stock_prices.RDS"))
    saveRDS(tuesdata$big_tech_companies, file = here("data", "tuesdata_big_tech_companies.RDS"))
}


big_tech_stock_prices <- readRDS(here("data", "tuesdata_big_tech_stock_prices.RDS"))
big_tech_companies <- readRDS(here("data", "tuesdata_big_tech_companies.RDS"))

## data wrangling
stock_yr_m=big_tech_stock_prices %>%
  full_join(big_tech_companies, by="stock_symbol")%>%
  mutate(year_month = format(date, "%Y-%m"))%>%
  group_by(year_month, stock_symbol)%>%
  filter(date==min(date)|date==max(date))%>%
  print()

## keep the open price of the first day of the month 
stock_open=stock_yr_m%>%
  group_by(stock_symbol,year_month)%>%
  filter(date==min(date))%>%
  select(stock_symbol, open, year_month, company)
## keep the adj price of the last day of the month 
stock_adj_close=stock_yr_m%>%
  group_by(stock_symbol,year_month)%>%
  filter(date==max(date))%>%
  select(stock_symbol, adj_close, year_month, company)

## simple rate of return 
stock_month=stock_open%>%
  left_join(stock_adj_close, by=c("stock_symbol", "year_month", "company"))%>%
  mutate(date=ym(year_month))%>%
  mutate(rate_of_return=(adj_close-open)/open)

stock_month%>%
ggplot(aes(x=date, y=adj_close))+
  geom_line(aes(x=date, y=adj_close, color="adj_close"))+
  geom_line(aes(x=date, y=open, color="open"))+
  facet_wrap(~stock_symbol,scales = "fixed", nrow=4)+
  scale_color_manual(values = c("adj_close" = "red", "open"= "blue"))+
  labs(title = "The open and adjust close price for 14 tech company from 2010-2023 ",
       subtitle = "Overall,most companies have pretty close open price and adj close price, the blue line and red line are overlapping. IBM is the only exception, the adj_close price mostly lower than the open price", x="date", y="stock price(dollar)", caption = "Chuyi Sun")

par(mfrow = c(1, 2))
stock_month%>%
  filter(stock_symbol=="IBM")%>%
  ggplot(aes(x=stock_symbol, y=open))+
  geom_boxplot()+
  labs(title = "The open price of IBM in each month(2010-2023)",
       subtitle = "The adj price is roughly evenly distributed aromg 143, the range is 101-202", x="date", y="stock price(dollar)", caption = "Chuyi Sun")+
  theme_bw()

stock_month%>%
  filter(stock_symbol=="IBM")%>%
  ggplot(aes(x=stock_symbol, y=adj_close))+
  geom_boxplot() +
  labs(title = "The adj price of IBM in each month(2010-2023)",
       subtitle = "The adj price is evenly distributed around 112, the range is 73 -149. ", x="date", y="stock price(dollar)", caption = "Chuyi Sun")+
  theme_bw()

IBM_rate_ci=make_ci_class(stock_month[stock_month$stock_symbol=="IBM",]$rate_of_return)
IBM_rate_ci
calculate_CI(IBM_rate_ci)
## does not include zero, there is a significant change in the rate
IBM_adj_close=make_ci_class(stock_month[stock_month$stock_symbol=="IBM",]$adj_close)
IBM_adj_close
calculate_CI(IBM_adj_close)

stock_month%>%
  group_by(stock_symbol)%>%
  summarise(rate_of_return_ci_low=calculate_CI(make_ci_class(rate_of_return))[1],
            rate_of_return_ci_upr=calculate_CI(make_ci_class(rate_of_return))[2])

stock_month%>%
  filter(date==ymd("2020-06-01"))%>%
  ggplot(aes(x=fct_reorder(.f=stock_symbol,.x=rate_of_return), y=rate_of_return, fill=stock_symbol))+
  geom_bar(stat="identity")+
  labs(title = "The rate of return for each company in mid 2020",
       subtitle = "More than half company have postive rate of return. Tesla has the largest rate of return and IBM has the lowest rate of return.", caption = "Chuyi Sun")+
  theme(axis.text.x = element_text(size = 9))


```
## Analysis summary  

When looking at the open price of the stock at the beginning of the month, and the adjusted close price at the end of the month, we can find the IBM have obvious difference in the plot. 
Then, we calculate the rate of the return to see the performance of the IBM from 2010-2023, and get the 95% confidence interval, the confidence interval shows there is a significant change in the rate of return, and the confidence interval is negative. 
From the overall trend, see a lot company have a clear increase in 2020, so we plot the rate of return of 14 company to see the pereformance.


## Function used 

##### Dplyr  

* full_join
* mutate
* group_by
* filter
* select
* left_join
* summarise

##### lubridate  

* ym
* ymd 

##### forcats  

* fct_reorder

##### ggplot2  

* geom_line
* geom_boxplot
* geom_bar

